---
title: "Homework 3"
author: "Jeanette Shekelle"
date: "10/9/2018"
output: github_document
---

```{r}
library(tidyverse)

```

```{r}
library(p8105.datasets)
```

## Problem 1

Data Import

```{r}
data("brfss_smart2010")
```



Data Cleaning:
* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Excellent” to “Poor”

```{r}
brfss_smart2010 = 
  brfss_smart2010 %>% 
   janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>% 
  mutate(response_level = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor")))

```



Answering the questions
* In 2002, which states were observed at 7 locations?

```{r}
brfss_smart2010 %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  distinct(locationdesc) %>% 
  count() %>% 
  filter(n == 7)
```

There are 3 states that were observed at 7 locations: CT, FL, and NC. 


* Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r}
library(ggplot2)
```


```{r}
brfss_smart2010 %>% 
  group_by(locationabbr) %>% 
  mutate(locations_per_state = count(distinct(locationdesc))) %>% 
  ggplot(aes(x = locationabbr, y = locations_per_state)) +
  geom_line() + geom_point() +
  theme_bw()

```

ggplot(brfss_smart2010, aes(x = locationdesc, y=, color=factor(ID))) +
  geom_line() + geom_point() +
  theme_bw()
```{r}
brfss_smart2010 %>% 
  
```


* Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
* For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

