---
title: "Homework 3"
author: "Jeanette Shekelle"
date: "10/9/2018"
output: github_document
---

Loading libraries

```{r}
library(tidyverse)
library(p8105.datasets)

```


## Problem 1

Data Import

```{r}
data("brfss_smart2010")
```



Data Cleaning:
* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Excellent” to “Poor”

```{r}
brfss_smart2010 = 
  brfss_smart2010 %>% 
   janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>% 
  mutate(response_level = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor")))

```



Answering the questions
* In 2002, which states were observed at 7 locations?

```{r}
brfss_smart2010 %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  distinct(locationdesc) %>% 
  count() %>% 
  filter(n == 7)
```

There are 3 states that were observed at 7 locations: CT, FL, and NC. 


* Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r}
library(ggplot2)
```


```{r}
brfss_smart2010 %>% 
  group_by(locationabbr, year) %>% 
  distinct(locationdesc) %>% 
  summarize(locations_per_state = n()) %>% 
  ggplot(aes(x = year, y = locations_per_state, color = locationabbr)) +
  geom_line() + theme_bw() + 
  viridis::scale_color_viridis(name = "States", discrete = TRUE) +
  labs (title = "Number of distinct locations in each state 2002-2010")

```

It is really hard to distinguish the plots of different states. Many colors look basically the same. I tried changing the color scale but it is still hard to tell. We need 50 colors that all look very different from one another and I am not sure that is possible. I don't think this is the best way to visualize this data because we can't tell what state had this huge spike in 2007 and 2010.  


* Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r}
brfss_smart2010 %>% 
  filter(year == 2002 | year == 2006 | year == 2010) %>% 
  filter(locationabbr == "NY") %>% 
  filter(response == "Excellent") %>% 
  group_by(year) %>% 
  summarise(mean_excellent = mean(data_value),
            sd_excellent = sd(data_value)) %>% 
  knitr::kable(digits = 2)
 
```



* For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r}
brfss_smart2010 %>% 
  group_by(year, locationabbr, response) %>% 
  summarize(mean_proportion = mean(data_value)) %>% 
  ggplot(aes(x = year, y = mean_proportion, color = locationabbr)) +
  geom_line() +
  facet_grid(~response)

```



## Problem 2

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

```{r}
library(p8105.datasets)
data("instacart") 
```


The instacart dataset has 15 variables and 1,384,617 observations. The key variables are the product name, aisle, and department because they tell us about the product. Other key variables are the the day of the week the product was ordered and the time of day. This information could be useful for planning product sales. 



### Some questions

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  distinct(aisle_id) %>% 
  nrow()
```

There are 134 aisles.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  count() %>% 
  min_rank(desc(n))
```

The most items are ordered from the following aisles: 

Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  group_by(aisle) %>%
  summarize(items_per_aisle = n()) %>% 
  ggplot(aes(x = aisle, y = items_per_aisle)) +
  geom_point()
```


Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
instacart %>% 
  filter(aisle == "baking ingredients" | 
           aisle == "dog food care" | 
           aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(most_popular = n())
  
```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_hour)
```



## Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
```


